---
title: "Prepare nphk1 dataset for publication on GBIF"
author: 
- Sanne Govaert
- Jim Casaer
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Load libraries

```{r prep packages}
library(here)
library(camtrapdp)
library(dplyr) 
library(rgbif)
library(lubridate)
library(EML)
library(zip)
```

## Read data

```{r camtrap data}
name_exportfolder <- "nphk-national-park-hoge-kempen-20260828104130"
file <- here::here("datasets", "nphk1", "data", "raw", name_exportfolder, "datapackage.json")
dataset <- camtrapdp::read_camtrapdp(file)
```

## Partial validation

Validation against the Camtrap DP specificaitons is only possible with [frictionless-py](https://camtrap-dp.tdwg.org/#validation).

Here we check for some common issues that need to be fixed in Agouti.

Load custom functions for validation, and set Agouti project id:

```{r}
source(here::here("functions", "validate.R"))
agouti_id <- "fed8eb80-9339-44ab-8801-de302be0357b"
```

### Validate deployments

Checks if any of these 4 required fields is empty:
- latitude
- longitude
- deploymentStart
- deploymentEnd

```{r}
validate_deployments(dataset, agouti_id)
```

### Validate observations

Checks for the following issues:
- eventStart or eventEnd is empty
- eventStart is before the start of the project
- count is 0
- count is NA if scientificName is not NA

```{r}
validate_observations(dataset, agouti_id)
```

### Check taxonomic scope

```{r}
taxa(dataset)
```


# Get data ready for publication

## Data

### Split dataset in NPHK1 and NPHK2

See https://github.com/inbo/camera-trap-datasets/issues/37

```{r}
dataset <-
  dataset |>
  filter_deployments(
    deploymentStart <= lubridate::as_date("2020-04-30")
  ) 
```


### Geospatial coordinates are rounded to 0.001 degrees

```{r round coordinates}
dataset <-
  dataset |> 
  camtrapdp::round_coordinates(digits = 3)
```

## Make GBIF-proof

GBIF does not accept negative values in bounding box values, so set negative values to NA.

``` {r}
camtrapdp::observations(dataset) <- 
  camtrapdp::observations(dataset) |>
  dplyr::mutate(
    bboxX = dplyr::if_else(.data$bboxX < 0, NA, bboxX),
    bboxY = dplyr::if_else(.data$bboxY < 0, NA, bboxY)
    ) |>
  dplyr::mutate(
    bboxX = dplyr::if_else(.data$bboxX > 1, NA, bboxX),
    bboxY = dplyr::if_else(.data$bboxY > 1, NA, bboxY)
    )
```
 

## Metadata

Add:

- title
- license
- abstract (description)
- keywords
- references
- project path

Adjust:

- contributors
- image url

```{r add metadata}
# Basic metadata
dataset$id <- "https://doi.org/10.15468/suawqp"

dataset$title <- "NPHK1 - Systematic random camera trap observations in National Park Hoge Kempen (Belgium)"

dataset$licenses <- list(
  list(
    scope = "data", 
    name = "CC0-1.0", 
    path = "https://creativecommons.org/publicdomain/zero/1.0/"
    ), 
  list(
    scope = "media",
    name = "CC-BY-4.0",
    path = "http://creativecommons.org/licenses/by/4.0/"
    )
  )

dataset$description <- paste(
  paste0(
    "<emphasis>", dataset$title, "</emphasis> is a dataset published by the <ulink url=\"https://www.inbo.be/en\"><citetitle>Research Institute for Nature and Forest (INBO)</citetitle></ulink>. It contains camera trap data primarily collected to monitor the interaction between human activities (e.g. recreation, hunting) and the population dynamics, as well as geographical and temporal distribution of small and medium-sized mammals (and birds) in the National Park Hoge Kempen (NPHK). Observations of this dataset span a monitoring period of three years, i.e. May 2017 until May 2020. The data were collected through close colloration with Hasselt University. Most of the fieldwork and annotations were done by staff and students of Hasselt University."),
    
    "We used a systematic-random sampling design. We divided the NPHK into 40 compartments of approximately 1.5 km², and superimposed a grid of 300 m x 300 m on these compartments. Each month, we randomly selected one grid cell within each of the 40 compartments, and used their midpoints as the sampling locations to set up a single camera trap (Reconyx HC600 Hyperfire) at the nearest tree.
We did not use bait nor did we select for trails. We placed camera traps on a height of 50 cm facing north without altering the surroundings, unless some vegetation was directly in front of the camera. Cameras were set up to the fastest trigger speed for a burst of ten pictures without delay between triggers and with time-lapse pictures every 12 hours to control for malfunction. We deployed camera traps for three consecutive weeks and relocated them in the fourth week, which resulted in 13 deployments per year. For further descriptions of the study site and sampling protocol see (<ulink url=\"https://doi.org/10.1016/j.scitotenv.2020.140369\"><citetitle>Wevers et al. 2020</citetitle></ulink>) and (<ulink url=\"https://doi.org/10.1007/s10344-025-02013-3\"><citetitle>Indestege et al. 2025</citetitle></ulink>)." ,

"Images from the camera traps were uploaded to <ulink url=\"https://www.agouti.eu\"><citetitle>Agouti</citetitle></ulink> for data annotation. Issues with the dataset can be reported at <ulink url=\"https://github.com/inbo/camera-trap-datasets/issues\"><citetitle>https://github.com/inbo/camera-trap-datasets/issues</citetitle></ulink>.",

    "We have released this dataset to the public domain under a Creative Commons Zero waiver. We kindly ask users to follow the INBO norms for data use (<ulink url=\"https://www.inbo.be/en/norms-data-use\"><citetitle>https://www.inbo.be/en/norms-data-use</citetitle></ulink>) when using the data. If you have any questions regarding this dataset, don't hesitate to contact us via the contact information provided in the metadata or via opendata@inbo.be. Researchers can request more detailed geographic coordinates of the camera trap locations by contacting jim.casaer@inbo.be or natalie.beenaerts@uhasselt.be",

    "This dataset was collected using infrastructure managed by INBO and funded by the Research Foundation – Flanders (FWO) as part of Belgium’s contribution to LifeWatch.",
    sep = "\n"
  )

dataset$contributors <- list(
  list(   
    title = "Jim Casaer",
    email = "jim.casaer@inbo.be",
    path = "https://orcid.org/0000-0001-6788-5876",
    role = "principalInvestigator",
    organization = "Research Institute for Nature and Forest (INBO)"
  ),
  list(
    title = "Jolien Wevers",
    path = "https://orcid.org/0000-0002-8453-2199",
    role = "contributor"
    ),
  list(
    title = "Martijn Bollen",
    email = "martijn.bollen@inbo.be",
    path = "https://orcid.org/0000-0002-1621-6258",
    role = "contributor",
    organization = "Research Institute for Nature and Forest (INBO)"
    ),
  list(
    title = "Peter Desmet",
    email = "peter.desmet@inbo.be",
    path = "https://orcid.org/0000-0002-8442-8025",
    role = "contributor",
    organization = "Research Institute for Nature and Forest (INBO)"
  ),
  list(
    title = "Sanne Govaert",
    email = "sanne.govaert@inbo.be",
    path = "https://orcid.org/0000-0002-8939-1305",
    role = "contributor",
    organization = "Research Institute for Nature and Forest (INBO)"
  ),
list(
    title = "Tanja Milotic",
    email = "tanja.milotic@inbo.be",
    path = "https://orcid.org/0000-0002-3129-6196",
    role = "contributor",
    organization = "Research Institute for Nature and Forest (INBO)"
  ),
  list(
    title = "Siebe Indestege",
    email = "siebe.indestege@uhasselt.be",
    path = "https://orcid.org/0009-0001-0901-6338",
    role = "contributor",
    organization = "CMK, Hasselt University"
    ),
  list(
    title = "Wim Kuypers",
    email = "wim.kuypers@uhasselt.be",
    path = "https://orcid.org/0000-0001-7264-4073",
    role = "contributor",
    organization = "CMK, Hasselt University"
    ),
  list(
    title = "Imke Tomsin",
    email = "imke.tomsin@uhasselt.be",
    path = "https://orcid.org/0009-0009-4894-1487",
    role = "contributor",
    organization = "CMK, Hasselt University"
    ),
  list(title = "Natalie Beenaerts",
    email = "natalie.beenaerts@uhasselt.be",
    path = "https://orcid.org/0000-0001-5655-5943",
    role = "principalInvestigator",
    organization = "CMK, Hasselt University"
    ),
  list(
    title = "Research Institute for Nature and Forest (INBO)",
    path = "https://www.inbo.be",
    role = "rightsHolder"
  )
)

# Fix mistake in image url
dataset$image <- "https://api.agouti.eu/uploads/project-images/fed8eb80-9339-44ab-8801-de302be0357b.JPG"

# Keywords
dataset$keywords <- c("camera traps", "alien species", "Flanders", "Belgium", "Mammalia", "birds", "Lifewatch", "invasive alien species", "European observatory of wildlife","Nationaal Park Hoge Kempen")

# References
dataset$references <- c(
  "Wevers J, Fattebert J, Casaer J, Artois T, and Beenaerts N (2020). Trading fear for food in the Anthropocene: How ungulates cope with human disturbance in a multi-use, suburban ecosystem. Science of the Total Environment 740: 140369. https://doi.org/10.1016/j.scitotenv.2020.140369",
  "Indestege S, Bollen M, Casaer J, and Beenaerts N (2025). Do long-term camera trap sampling designs matter to estimate diel activity? European Journal Of Wildlife Research 71:132. https://doi.org/10.1007/s10344-025-02013-3"
)

# Add project path
dataset$project$path <- "https://agouti.eu/explore/project/fed8eb80-9339-44ab-8801-de302be0357b"

# Add project acronym
dataset$project$acronym <- "NPHK"
```


# Write Camera Trap Data Package

```{r write camtrap DP}
dir_processed <- here::here("datasets", "nphk1", "data", "processed")
camtrapdp::write_camtrapdp(dataset, dir_processed)

# Zip these files
zipfile <-  paste0(dir_processed, "/datapackage.zip")
files2zip <- c(
  "datapackage.json",
  "deployments.csv",
  "media.csv",
  "observations.csv"
)
zip::zip(zipfile, files = files2zip, root = dir_processed)
```
