---
title: "Prepare Agouti project Drongengoed (Camtrap DP) for publication on GBIF"
author: 
- Sanne Govaert
- Jim Casaer
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Load libraries

```{r prep packages}
library(here)
library(camtrapdp)
library(dplyr) 
library(rgbif)
library(lubridate)
library(EML)
library(zip)
```

## Read data

```{r camtrap data}
name_exportfolder <- "drongengoed-adaptief-grofwildbeheer-20260826120331"
file <- here::here("datasets", "drongengoed", "data", "raw", name_exportfolder, "datapackage.json")
dataset <- camtrapdp::read_camtrapdp(file)
```

# Get data ready for publication

## Data

### Geospatial coordinates are rounded to 0.001 degrees

```{r round coordinates}
dataset <-
  dataset |> 
  camtrapdp::round_coordinates(digits = 3)
```

## Make GBIF-proof

GBIF does not accept negative values in bounding box values, so set negative values to NA.

``` {r}
camtrapdp::observations(dataset) <- 
  camtrapdp::observations(dataset) |>
  dplyr::mutate(
    bboxX = dplyr::if_else(.data$bboxX < 0, NA, bboxX),
    bboxY = dplyr::if_else(.data$bboxY < 0, NA, bboxY)
    ) |>
  dplyr::mutate(
    bboxX = dplyr::if_else(.data$bboxX > 1, NA, bboxX),
    bboxY = dplyr::if_else(.data$bboxY > 1, NA, bboxY)
    )
```
 

## Metadata

Add:

- title
- license
- abstract (description)
- keywords
- references
- project path

Adjust:

- contributors
- image url

```{r add metadata}
# Basic metadata
dataset$title <- "DRONGENGOED - "

dataset$licenses <- list(
  list(
    scope = "data", 
    name = "CC0-1.0", 
    path = "https://creativecommons.org/publicdomain/zero/1.0/"
    ), 
  list(
    scope = "media",
    name = "CC-BY-4.0",
    path = "http://creativecommons.org/licenses/by/4.0/"
    )
  )

dataset$description <- paste(
  paste0(
    "<emphasis>", dataset$title, "</emphasis> is a dataset published by the <ulink url=\"https://www.inbo.be/en\"><citetitle>Research Institute for Nature and Forest (INBO)</citetitle></ulink>. It contains camera trap data ..."),
    
    "We used a ..." ,

"Images from the camera traps were uploaded to <ulink url=\"https://www.agouti.eu\"><citetitle>Agouti</citetitle></ulink> for data annotation. Issues with the dataset can be reported at <ulink url=\"https://github.com/inbo/camera-trap-datasets/issues\"><citetitle>https://github.com/inbo/camera-trap-datasets/issues</citetitle></ulink>.",

    "We have released this dataset to the public domain under a Creative Commons Zero waiver. We kindly ask users to follow the INBO norms for data use (<ulink url=\"https://www.inbo.be/en/norms-data-use\"><citetitle>https://www.inbo.be/en/norms-data-use</citetitle></ulink>) when using the data. If you have any questions regarding this dataset, don't hesitate to contact us via the contact information provided in the metadata or via opendata@inbo.be. Researchers can request more detailed geographic coordinates of the camera trap locations by contacting jim.casaer@inbo.be.",

    "This dataset was collected using infrastructure managed by INBO and funded by the Research Foundation – Flanders (FWO) as part of Belgium’s contribution to LifeWatch.",
    sep = "\n"
  )

dataset$contributors <- list(
  list(   
    title = "Jim Casaer",
    email = "jim.casaer@inbo.be",
    path = "https://orcid.org/0000-0001-6788-5876",
    role = "principalInvestigator",
    organization = "Research Institute for Nature and Forest (INBO)"
  ),
  list(
    title = "Sanne Govaert",
    email = "sanne.govaert@inbo.be",
    path = "https://orcid.org/0000-0002-8939-1305",
    role = "contributor",
    organization = "Research Institute for Nature and Forest (INBO)"
  ),
  list(
    title = "Research Institute for Nature and Forest (INBO)",
    path = "https://www.inbo.be",
    role = "rightsHolder"
  )
)

# Fix mistake in image url
dataset$image <- "https://api.agouti.eu/uploads/project-images/20230811080824-20230522191931-RCNX1406%20(1).JPG"

# Keywords
dataset$keywords <- c("")

# References
# dataset$references <- ""

# Add project path
dataset$project$path <- "https://agouti.eu/explore/project/e10e6b2b-78fc-44ad-9d66-c4d0b1dd889e"

# Add project acronym
dataset$project$acronym <- "Drongengoed"
```


# Write Camera Trap Data Package

```{r write camtrap DP}
dir_processed <- here::here("datasets", "drongengoed", "data", "processed")
camtrapdp::write_camtrapdp(dataset, dir_processed)

# Zip these files
zipfile <-  paste0(dir_processed, "/datapackage.zip")
files2zip <- c(
  "datapackage.json",
  "deployments.csv",
  "media.csv",
  "observations.csv"
)
zip::zip(zipfile, files = files2zip, root = dir_processed)
```
