---
title: "Prepare Agouti project GMU8_LEUVEN (Camtrap DP) for publication on GBIF"
author: 
- Sanne Govaert
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Load libraries

```{r prep packages}
library(here)
library(camtrapdp)
library(dplyr) 
library(rgbif)
library(lubridate)
library(EML)
library(zip)
```

## Read data

```{r camtrap data}
name_exportfolder <- "gmu8-monitoring-faunabeheerzone-8-20250515135034"
file <- here::here("datasets", "GMU8_LEUVEN", "data", "raw", name_exportfolder, "datapackage.json")
data <- camtrapdp::read_camtrapdp(file)
```

# Get data ready for publication

## Data

### Geospatial coordinates are rounded to 0.001 degrees

```{r round coordinates}
rounded_data <-
  data %>% 
  camtrapdp::round_coordinates(digits = 3)
```

### Set embargo

Set an embargo on the last year. Before this date, all data is public. After this date, only observations of exotic species are kept.

First, construct a vector with the alien mammals and birds of Belgium.

```{r}
taxa_mammals <- rgbif::name_usage(datasetKey = "9a52d8bf-864a-4abb-95ba-319c4edfca8d", limit = 99999)$data
taxa_birds <- rgbif::name_usage(datasetKey = "e1c3be64-2799-4342-8312-49d076993132", limit = 99999)$data

exotic_species <- c(
  dplyr::pull(taxa_mammals, species),
  dplyr::pull(taxa_birds, species)
) %>% 
  na.omit()
```


```{r}
today <- Sys.Date()
embargo_date <- today %m-% months(12)
embargo_data <- 
  rounded_data %>% 
  camtrapdp::filter_observations(
    eventStart <= embargo_date
    |
    scientificName %in% exotic_species
  )
```

Check which species are maintained after the embargo data:

```{r}
embargo_data %>%
  camtrapdp::filter_observations(eventStart >= embargo_date) %>% 
  camtrapdp::taxa()
```


### Privacy settings

Hide the media of dogs for privacy reasons.

```{r remove media of dogs}
mediaID_dogs <- embargo_data %>% 
  camtrapdp::filter_observations(scientificName %in% c("Canis lupus familiaris")) %>% 
  camtrapdp::media() %>% 
  dplyr::pull(mediaID)

camtrapdp::media(embargo_data) <-
  rounded_data %>%
  camtrapdp::media() %>%
  dplyr::mutate(
    filePath = dplyr::if_else(
      mediaID %in% mediaID_dogs,
      "private",
      filePath,
      missing = filePath
    ),
    filePublic = dplyr::if_else(
      mediaID %in% mediaID_dogs,
      FALSE,
      filePublic,
      missing = filePublic
    )
  )
```

## Make GBIF-proof

GBIF does not accept negative values in bounding box values, so set negative values to NA.

```{r}
gmu8_leuven <- embargo_data
camtrapdp::observations(gmu8_leuven) <- 
  camtrapdp::observations(gmu8_leuven) %>%
  dplyr::mutate(
    bboxX = dplyr::if_else(.data$bboxX < 0, NA, bboxX),
    bboxY = dplyr::if_else(.data$bboxY < 0, NA, bboxY)
    ) %>%
  dplyr::mutate(
    bboxX = dplyr::if_else(.data$bboxX > 1, NA, bboxX),
    bboxY = dplyr::if_else(.data$bboxY > 1, NA, bboxY)
    )
```
 

## Metadata

Add:

- title
- license
- abstract (description)
- keywords
- references
- project path

Adjust:

- contributors
- image url

```{r add metadata}
# Basic metadata
gmu8_leuven$title <- "GMU8_LEUVEN - Camera trap observations in natural habitats south of Leuven (Belgium)" 

gmu8_leuven$licenses <- list(
  list(
    scope = "data", 
    name = "CC0-1.0", 
    path = "https://creativecommons.org/publicdomain/zero/1.0/"
    ), 
  list(
    scope = "media",
    name = "CC-BY-4.0",
    path = "http://creativecommons.org/licenses/by/4.0/"
    )
  )

gmu8_leuven$description <- c(
  paste0(
    "<p><b>", gmu8_leuven$title, "</b> is a dataset published by the <a href=\"https://www.inbo.be/en\">Research Institute of Nature and Forest (INBO)</a>. It contains camera trap data primarily collected to monitor the recolonization of wild boar in natural areas south of Leuven. However, the dataset also includes all recorded observations of mammals and birds captured during the monitoring period.</p>",
    "<p>This monitoring was conducted in close collaboration with hunters and members of local nature conservation associations active in the area. Several forest and nature areas south of Leuven were sampled, with the main forest complexes being <b>Meerdaalwoud</b>, <b>Heverleebos</b>, and <b>Egenhovenbos</b>. Additionally, camera traps were placed in the valleys of the <b>Dijle</b>, <b>Laan</b>, and <b>Yse</b> rivers. Each camera trap was installed facing north at a height of approximately 50 cm. No bait was used during the monitoring. The sampling design was based on a grid of 250 x 250 meter cells. Camera locations were determined by randomly selecting a subset of these grid cells, and the coordinates used for placement corresponded to the center point of each selected cell. A total of <b>32 camera traps</b> were used. These were relocated monthly, with each selected grid cell being monitored once during the winter and once during the summer. All images captured by the camera traps were uploaded to <a href=\"https://www.agouti.eu\">Agouti</a> for data annotation. Issues with the dataset can be reported at <a href=\"https://github.com/inbo/camera-trap-datasets/issues\">https://github.com/inbo/camera-trap-datasets/issues</a>.</p>",
    "<p>We have released this dataset to the public domain under a Creative Commons Zero waiver. We kindly ask users to follow the INBO norms for data use (<a href=\"https://www.inbo.be/en/norms-data-use\">https://www.inbo.be/en/norms-data-use</a>) when using the data. If you have any questions regarding this dataset, don't hesitate to contact us via the contact information provided in the metadata or via opendata@inbo.be. Researchers can request more detailed geographic coordinates of the camera trap locations by contacting jim.casaer@inbo.be.</p>",
    "<p>This dataset was collected using infrastructure managed by INBO and funded by the <b>Research Foundation – Flanders (FWO)</b> as part of Belgium’s contribution to <b>LifeWatch</b>.</p>"
  )
)

gmu8_leuven$contributors <- list(
  list(
    title = "Jim Casaer",
    email = "jim.casaer@inbo.be",
    path = "https://orcid.org/0000-0001-6788-5876",
    role = "principalInvestigator",
    organization = "Research Institute for Nature and Forest (INBO)"
  ),
  list(
    title = "Niko Boone",
    email = "niko.boone@inbo.be",
    path = " https://orcid.org/0000-0001-7334-2819 ",
    role = "principalInvestigator",
    organization = "Research Institute for Nature and Forest (INBO)"
  ),
  list(
    title = "Jan Vercammen",
    email = "jan.vercammen@inbo.be",
    role = "contributor",
    organization = "Research Institute for Nature and Forest (INBO)"
  ),
  list(
    title = "Sander Devisscher",
    email = "sander.devisscher@inbo.be",
    path = " https://orcid.org/0000-0003-2015-5731",
    role = "contributor",
    organization = "Research Institute for Nature and Forest (INBO)"
  ),
  list(
    title = "Lynn Pallemaerts",
    email = "lynn.pallemaerts@inbo.be",
    path = "https://orcid.org/0000-0002-5034-5416",
    role = "contributor",
    organization = "Research Institute for Nature and Forest (INBO)"
  ),
  list(
    title = "Anneleen Rutten",
    email = "anneleen.rutten@inbo.be",
    path = "https://orcid.org/0000-0002-4025-5265",
    role = "contributor",
    organization = "Research Institute for Nature and Forest (INBO)"
  ),
  list(
    title = "Martijn Bollen",
    email = "martijn.bollen@inbo.be",
    path = "https://orcid.org/0000-0002-1621-6258",
    role = "contributor",
    organization = "Research Institute for Nature and Forest (INBO)"
  ),
  list(
    title = "Peter Desmet",
    email = "peter.desmet@inbo.be",
    path = "https://orcid.org/0000-0002-8442-8025",
    role = "contributor",
    organization = "Research Institute for Nature and Forest (INBO)"
  ),
  list(
    title = "Sanne Govaert",
    email = "sanne.govaert@inbo.be",
    path = "https://orcid.org/0000-0002-8939-1305",
    role = "contributor",
    organization = "Research Institute for Nature and Forest (INBO)"
  ),
  list(
    title = "Research Institute for Nature and Forest (INBO)",
    path = "https://www.inbo.be",
    role = "rightsHolder"
  )
)

# Fix mistake in image url
gmu8_leuven$image <- "https://api.agouti.eu/uploads/project-images/20180430144018-IMG_0612.JPG"

# Keywords
gmu8_leuven$keywords <- c("camera traps", "alien species", "Flanders, Belgium", "Mammalia", "birds", "Lifewatch", "invasive alien species", "European observatory of wildlife","Nationaal Park Brabantse Wouden")

# References
gmu8_leuven$references <- NULL

# Add project path
gmu8_leuven$project$path <- "https://agouti.eu/#/explore/project/91e1e104-a296-461e-af79-eeb6cb3e6c7f"
```


# Write Camera Trap Data Package

```{r write camtrap DP}
dir_processed <- here::here("datasets", "GMU8_LEUVEN", "data", "processed")
camtrapdp::write_camtrapdp(gmu8_leuven, dir_processed)

# Zip these files
zipfile <-  paste0(dir_processed, "/datapackage.zip")
files2zip <- c(
  "datapackage.json",
  "deployments.csv",
  "media.csv",
  "observations.csv"
)
zip::zip(zipfile, files = files2zip, root = dir_processed)
```

